---
title: "Glaucoma staging study"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r packages, results='hide'}
.packages <- c("tidyverse", "FactoMineR", "factoextra")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

```{r data-loading, results='hide'}
load("data/data-rims-preprocessed.RData")
df <- df.rims
# Set of variables
varBMO <- startsWith(colnames(df),"BMO")
varBMO[11] <- F
varBMO <- colnames(df)[varBMO]
var3.5 <- startsWith(colnames(df),"RNFL3.5")
var3.5 <- colnames(df)[var3.5]
var4.1 <- startsWith(colnames(df),"RNFL4.1")
var4.1 <- colnames(df)[var4.1]
var4.7 <- startsWith(colnames(df),"RNFL4.7")
var4.7 <- colnames(df)[var4.7]
varG <- endsWith(colnames(df), ".G")
varG <- colnames(df[varG])
varTI <- endsWith(colnames(df), ".TI")
varTI <- colnames(df[varTI])
varT <- endsWith(colnames(df), ".T")
varT <- colnames(df[varT])
varTS <- endsWith(colnames(df), ".TS")
varTS <- colnames(df[varTS])
varNS <- endsWith(colnames(df), ".NS")
varNS <- colnames(df[varNS])
varN <- endsWith(colnames(df), ".N")
varN <- colnames(df[varN])
varNI <- endsWith(colnames(df), ".NI")
varNI <- colnames(df[varNI])
varRims <- c(varBMO, var3.5, var4.1, var4.7)
varSectors <- c(varG, varTI, varT, varTS, varNS, varN, varNI)

df <- df %>% 
  # Variables selection
  select(c("Id", "Gender", "Age", "Glaucoma", "Eye", "BMO.Area", all_of(varRims))) %>%
  # Filter left eyes
  filter(Eye == "L") %>%
  # Select complete cases
  na.omit()

# Data frame of glaucoma eyes
df.glaucoma <- df %>% filter(Glaucoma == "Y")
```

# Number of clusters

```{r number-clusters}
fviz_nbclust(df.glaucoma[, varRims], kmeans, method = "wss") + 
  xlab("Number of clusters") + 
  ylab("Intra-groups sum of squares") + 
  ggtitle("Reduction of intra-group variability according to the number of clusters")
ggsave("img/reduction-variability-clusters.pdf", width = 7, height = 5)
```

According to the elbow criteria we decided to create 4 clusters, since the reduction in the intra-groups variability is not important for more than 4 groups.

# K-means clusters

```{r k-means-clusters}
# K-means clusters
seed = 123
# Variables considered
vars = varRims
# Number of clusters
n = 4
labels <- c("I", "II", "III", "IV")
# Color palette
# colfunc <- colorRampPalette(c("#FD8D3C", "#800026"))
#palette.healty <- colorRamps::green2red(5)
palette.healthy <- c("#00FF00", "#3FBF00", "#7F7F00", "#BF3F00", "#FF0000")
palette <- palette.healthy[-1]
# palette.healthy <- c("#00BFC4", palette)
# Shapes
shapes <- c(1, 3, 15, 17, 16)
# k-means
# C <- chol( var(df[df$Glaucoma=="Y", vars]) )
# y <- as.matrix(df[df$Glaucoma=="Y", vars]) %*% solve(C)
# clusters <- kmeans(y, centers = n, nstart = 25)
# centers <- clusters$centers[order(clusters$centers[,1], decreasing = T),]
# clusters <- kmeans(y, centers = centers)

clusters <- kmeans(df[df$Glaucoma=="Y", vars], centers = n, nstart = 25)
centers <- clusters$centers[order(clusters$centers[,1], decreasing = T),]
clusters <- kmeans(df[df$Glaucoma=="Y", vars], centers = centers)
# Convert the cluster into a factor
clusters$cluster <- as.factor(clusters$cluster)
# Assign labels to factors
levels(clusters$cluster) <- labels
# Add cluster to data frame
df$Stage <- factor("Healthy", levels=c("Healthy", labels))
df[df$Glaucoma=="Y", "Stage"] <- clusters$cluster
# Save clusters dafa frame
df %>% select(Id, Stage) %>% write_csv("data/data-stages.csv")
```

## Map of glaucoma stages whithout healty eyes

```{r map-clusters-whithout-healthy}
# Principal components
result.pca <- PCA(df[, vars], scale.unit = F, graph=F) 
# Map of clusters (without healthy eyes)
fviz_pca_ind(
  result.pca,
  geom = c("point"),
  axes = c(1, 2),
  select.ind = list(name = df %>% rownames_to_column() %>% filter(Glaucoma == "Y") %>% .$rowname),
  col.ind = df$Stage,
  palette = palette,
  addEllipses = T,
  ellipse.type = "t",
  title = "Map of glaucoma stages on the first two principal components",
  legend.title = "Stage"
) +
  xlab("First principal component") +
  ylab("Second principal component") +
  scale_shape_manual(values = shapes[-1]) +
  #coord_flip(clip = "off") +
  theme(legend.position = "top", plot.title = element_text(hjust = 0.5))
ggsave("img/map-clusters-whithout-healthy.pdf")
```

As can be seen in this chart, the four clusters can be separated almost perfectly along the x-axis of the first principal component. 

## Map of glaucoma stages whith healty eyes

```{r map-clusters-whith-healthy}
# Map of clusters (with healthy eyes)
fviz_pca_ind(
  result.pca,
  geom = c("point"),
  axes = c(1, 2),
  col.ind = df$Stage,
  palette = palette.healthy,
  addEllipses = T,
  ellipse.type = "t",
  title = "Map of glaucoma stages, with healthy eyes, on the first two principal components",
  legend.title = "Stage"
) +
  xlab("First principal component") +
  ylab("Second principal component") +
  scale_shape_manual(values = shapes) +
  theme(legend.position = "top", plot.title = element_text(hjust = 0.5))
ggsave("img/map-clusters-with-healthy.pdf")
```

## Distribution of the first principal component by glaucoma stages

```{r clusters-distribution}
# Add principal components coordinates to a data frame
df.pca <- as.data.frame(result.pca$ind$coord)
# Add cluster to principal components data frame
df.pca$Stage <- df$Stage
ggplot(df.pca, aes(x = Dim.1)) +
  geom_density(aes(fill = Stage), colour = I("white"), position = "identity", alpha = .5) + 
  ggtitle("Distribution of the first principal component according to glaucoma stages") +
  xlab("First principal component") +
  ylab("Probability density") +
  scale_fill_manual(values = palette.healthy) +
  theme(legend.position = "top", plot.title = element_text(hjust = 0.5))
ggsave("img/clusters-distributions.pdf", width = 7, height = 5)
```

## Confidence intervals of means by glaucoma stages

```{r confidence-intervals-means-by-stages}
# Means of variables by clusters
df.long <- df %>% select(Stage, vars) %>% pivot_longer(-Stage, names_to = "Sector") 

means <- df.long %>% group_by(Stage, Sector) %>% 
  summarise(n = n(), media = mean(value, na.rm = T), sd = sd(value, na.rm = T))  %>%
  mutate(se = sd/sqrt(n), lower.ci = media-qt(1-(0.05/2), n-1)*se, upper.ci = media+qt(1-(0.05/2), n-1)*se)

# Plot of means by clusters
ggplot(means, aes(x = Sector, y = media, colour = Stage, shape = Stage)) +
  geom_pointrange(aes(ymin = lower.ci, ymax = upper.ci), position = position_dodge(width = 0.5)) +
  ggtitle("Confidence intervals for the means of BMO and RNF rims sectors by stages") +
  scale_colour_manual(values = palette.healthy) +
  scale_shape_manual(values = shapes) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5), plot.title = element_text(hjust = 0.5))
```





