---
title: "Glaucoma staging study"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r data-loading, results='hide'}
source("data-loading.R")
```

```{r packages, results='hide'}
.packages <- c("reshape2")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

# Correlation analysis

```{r correlation-matrix}
# Correlation matrix
cormat <- round(cor(data[, varRims], use="complete.obs"),2)
melted.cormat <- melt(cormat)
melted.cormat$Var1 <- factor(melted.cormat$Var1, levels=varSectors)
melted.cormat$Var2 <- factor(melted.cormat$Var2, levels=varSectors)
print(ggplot(data = melted.cormat, aes(x=Var1, y=Var2, fill=value, label=value)) + 
        geom_tile() +
        geom_text(size=1) +
        xlab('') + 
        ylab('') + 
        ggtitle("Correlation matrix among BMO and RFNL rims by sectors") + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)))
```

The correlation analysis showed a clear correlation pattern by sectors. There was a very strong correlation (r>0.9) among the thickness of the BMO and RNFL rims 3.5, 4.1 and 4.7 by sectors. There was also a strong correlation (r>0.75) among the sectors of the BMO, except the BMO.G, that presented a very strong correlation with the others sectors. On the other hand, there was a very weak correlation (r<0.2) between the T sector and the N, NI and NS sectors. The T sector was in fact, the less correlated with the rest. 

