---
title: "Glaucoma staging study"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r packages, results='hide'}
.packages <- c("DT")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

# Authors

- Alfonso Parra Blesa ([alfonsoparrablesa@gmail.com](mailto:alfonsoparrablesa@gmail.com))
- Alfredo Sánchez Alberca ([asalber@ceu.es](mailto:asalber@ceu.es))
- José J García-Medina ([josegarciam@yahoo.com](mailto:josegarciam@yahoo.com))

# Objetives

Develop a staging system for glaucoma and a predictive model to classify new patiens in the glaucoma stages using optical coherence tomography (OCT) data.

# Type of study

Analityc observational cross-sectional and comparative study.

# Inclusion criteria

Only left eyes were considered.

# Sample size

```{r data-loading}
source("data-loading.R")
```


Total number of eyes: **1001** (766 Healthy and 235 with Glaucoma).
```{r sample-size-groups}
# table.sizes <- table(data$Glaucoma)
# names(table.sizes) <- c("Healthy", "Glaucoma")
# print(table.sizes)
```

# Variables

```{r}
variables <- read.csv(file="data/variables.csv", encoding="UTF-8", header=T, sep=",")
datatable(variables, rownames = F, escape=F, class = 'display', options = list(pageLength = 10, dom="ltip", language = list(url = 'http://cdn.datatables.net/plug-ins/1.10.11/i18n/Spanish.json')))
variables.Num <- variables[variables$Type=="Num",]
variables.Factors <- variables[variables$Type=="Factor",]
```

# Data preprocessing

It is known that the thickness of the retinal nerve fiber layers depends on age and BMO area. To avoid this dependency a linear transformation provided by the OCT company was applied to every variable. This linear transformations was obtained fitting a linear regression model on a normative group of healthy eyes.

$$z_i=(x_i-\bar x-b_{xe}(e_i-\bar e)-b_{xa}(a_i-\bar a))/s_x$$
where:

- $x_i$ is the value of variable $x$ on eye $i$.
- $\bar x$ is the mean of $x$.
- $s_x$ is the standard deviation of $x$.
- $e_i$ is the age of individual $i$.
- $\bar e$ is the mean of the age in the normative database of healthy eyes.
- $b_{xe}$ is the slope of the regression line of variable $x$ on the age.
- $a_i$ is the BMO area of eye $i$.
- $\bar a$ is the mean of the BMO area in the normative database of healthy eyes.
- $b_{xa}$ is the slope of the regression line of variable $x$ on the BMO area.
- $z_i$ is the standardized value of variable $x$ on eye $i$.


# Glosary

- OAG: Glaucoma de ángulo abierto
- OCT: Optical coherence tomography
- BMO: Bruch Membrane Opening 