---
title: "Glaucoma staging study"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r data-loading, results='hide'}
source("data-loading.R")
```

```{r packages, results='hide'}
.packages <- c("FactoMineR", "factoextra")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

# Principal components analysis

```{r principal-components}
# Principal components healthy and glaucoma eyes
result.pca <- PCA(data[, c(varRims)], scale.unit = F, graph=F)
# eigenvalues <- get_eigenvalue(result.pca)
# Explained variability by principal components
fviz_eig(result.pca,
  main = "Explained variability by principal components",
  ylab = "% of total variability",
  xlab = "Principal components (dimensions)") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("img/explained-variability-principal-components.pdf")
# Correlation of variables with first principal component
fviz_cos2(result.pca, 
  choice = "var", 
  axes=1, 
  title="Correlation of variables with the first principal component",
  ) + 
  ylab("Coef. Determination r²") +
  theme(plot.title = element_text(hjust = 0.5))
# Contribution of variables to principal
fviz_contrib(result.pca, choice="var", axes=1, top = 10) +
  theme(plot.title = element_text(hjust = 0.5))
```

The principal components analysis shows that the first principal component explains almost 60% of the total variability, while the second one only 11%.

The variables most correlated with the first principal component are the global sector and the TI sector of BMO and RNFL rims.

```{r map-eyes-principal-components}
# Add coordinates of principal components to data frame
data.pca <- cbind(result.pca$ind$coord, data[, "Glaucoma", drop=F])
# Plot eyes on the two first principal components
fviz_pca_ind(
  result.pca,
  geom = c("point"),
  axes = c(1, 2),
  habillage = data.pca$Glaucoma,
  ddEllipses = T,
  ellipse.type = "t",
  title = "Map of healthy and glaucoma eyes on the first two principal components"
) +
  scale_shape_discrete(labels = c("Glaucoma", "Healthy")) +
  scale_color_discrete(labels = c("Glaucoma", "Healthy")) +
  theme(legend.position = "top", plot.title = element_text(hjust = 0.5))
```

The map shows a clear separation, with a small overlapping area, between healthy and glaucoma eyes along the firs principal component, while there is no separation at all along the second one.



