---
title: "Glaucoma staging study"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r packages, results='hide'}
.packages <- c("tidyverse", "irr")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

```{r data-loading, results='hide'}
data <- read_csv('data/visual-field.csv', col_types = cols(Stage = "f"))
data$Stage <- ordered(data$Stage, levels=levels(data$Stage))                 
```

```{r data-filtering, results='hide'}
data <- data %>%  # Elminar los casos sin datos en la variable Dm
  filter(!is.na(Dm)) %>% # Eliminar los casos con errores mayores del 16%
  filter(ERROR <= 16)
```

# Analysis of visual field 

## Boxplot of visual field by Glaucoma stages

```{r distribution}
palette.healthy <- c("#00FF00", "#3FBF00", "#7F7F00", "#BF3F00", "#FF0000")
palette <- palette.healthy[-1]

data %>% 
  ggplot(aes(x=Dm, fill=Stage)) +
  geom_boxplot() +
  scale_fill_manual(values = palette)
```

## Conficence intervals for the mean by Glaucoma stages

  ```{r confidence-intervals-glaucoma-stages}
  means <- data %>% 
    group_by(Stage) %>%
    summarise(n=n(), mean = mean(Dm), sd= sd(Dm)) %>%
    mutate(se = sd / sqrt(n), lo.ci = mean - qt(1 - (0.05 / 2), n - 1) * se, up.ci = mean + qt(1 - (0.05 / 2), n - 1) * se)
  kable(means)
  ```

```{r confidence-intervals-plot}
# Color palette
ggplot(means, aes(x=Stage, y = mean, colour = Stage, shape = Stage)) +
  geom_pointrange(aes(ymin = lo.ci, ymax = up.ci), position = position_dodge(width = 0.5)) +
  ggtitle("Confidence intervals for the means of the visual field by Glaucoma Stages") +
  scale_colour_manual(values = palette) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5), plot.title = element_text(hjust = 0.5))
```

## Visual field stages

We define 4 stages using the visual field. 

```{r k-means-clusters}
# K-means clusters
seed = 123
# Variables considered
vars = "Dm"
# Número de clusters
n = 4
labels <- c("I", "II", "III", "IV")
# Color palette
# colfunc <- colorRampPalette(c("#FD8D3C", "#800026"))
#palette.healty <- colorRamps::green2red(5)
palette.healthy <- c("#00FF00", "#3FBF00", "#7F7F00", "#BF3F00", "#FF0000")
palette <- palette.healthy[-1]
# palette.healthy <- c("#00BFC4", palette)
# Shapes
shapes <- c(1, 3, 15, 17, 16)
# k-means
# C <- chol( var(data[data$Glaucoma=="Y", vars]) )
# y <- as.matrix(data[data$Glaucoma=="Y", vars]) %*% solve(C)
# clusters <- kmeans(y, centers = n, nstart = 25)
# centers <- clusters$centers[order(clusters$centers[,1], decreasing = T),]
# clusters <- kmeans(y, centers = centers)

clusters <- kmeans(data[, vars], centers = n, nstart = 25)
centers <- clusters$centers[order(clusters$centers[,1], decreasing = T),]
clusters <- kmeans(data[, vars], centers = centers)
# Convert the cluster into a factor
clusters$cluster <- as.factor(clusters$cluster)
# Assign labels to factors
levels(clusters$cluster) <- labels
# Add cluster to data frame
data[, "Stage.VF"] <- ordered(clusters$cluster)
# data %>% ggplot(aes(x=Dm, y=Stage.VF, colour=Stage.VF)) + geom_point()
```


```{r confidence-intervals-visual-field-stages}
means <- data %>% 
  group_by(Stage.VF) %>%
  summarise(n=n(), mean = mean(Dm), sd= sd(Dm)) %>%
  mutate(se = sd / sqrt(n), lo.ci = mean - qt(1 - (0.05 / 2), n - 1) * se, up.ci = mean + qt(1 - (0.05 / 2), n - 1) * se)
kable(means)
```

## Concordance analysis

We compute the Kendall's coefficient of concordance to check the agreement between the Glaucoma stages and the visual field stages.

```{r Kendalls-coefficient}
kendall(data[, c("Stage", "Stage.VF")], correct = T)
```
As $W=0.831$ there is a significant strong concordance between both rankings. 


